Extreme heterogeneity of influenza virus infection in single cells

### 2018-01-14 paper/ResponseToReviews.tex

\section*{MAJOR COMMENTS}

1. There are some obvious sources of variability that haven't been suitably discussed.

- The first is variation in the timing of the early stages of infection. No steps appear to have been taken to synchronize infection, or to limit secondary spread of the virus within the cultures. Could heterogeneity in the timing of binding/entry/fusion/trafficking explain a lot of the variation observed?

{\color{black}
This is an important point.
We did not synchronize infection, although for our 8-hour samples, we washed the cells one hour after infection for one sample (8hr) but not the other (8hr-2) in order to help evaluate the importance of infection timing.
The heterogeneity was similar between these two samples, as demonstrated by the new quantification method (Gini coefficients) suggested by the reviewers in a later comment.

We have also performed new experiments where we infected cells with synchronized infections by pre-binding virus on ice, washing it away, and then raising the temperature (this is a previously described procedure in the influenza virus literature).
Analysis of this experiment by flow cytometry staining is shown in the new Figure 4-Figure supplement 4, and shows that these synchronized infections look very similar to ones that are neither synchronized nor washed in terms of viral protein expression.

Finally, the barcoded viruses (Figure 4B) provide strong evidence against secondary infection, since secondary infection would tend to lead to mixed-barcode cells because during secondary infections there would be many of progeny viruses of both barcodes.

In addition to adding Figure 4-Figure supplement 4, we have added the following text related to this point:

\begin{quote}
\textsl{
One possible source of heterogeneity in the amount of viral mRNA per cell is variability in the timing of infection.
If some cells are infected earlier in the experiment than others, then they might have substantially more viral mRNA.
However, several lines of evidence indicate that this is not the major cause of heterogeneity across cells.
First, the sample for which the infection inoculum was never removed (8hr-2) only shows slightly more heterogeneity than samples for which the inoculum was washed away after one hour (Figure 4E, Figure 4-Figure~supplement~3), despite the fact that the potential time window for infection is much longer in the former sample.
Second, in an independent experiment, we performed completely synchronized infections by pre-binding virus to cells on ice and then washing away unbound virus before bringing the cells to 37$^{\circ}$C.
As shown in Figure 4-Figure~supplement 4, flow cytometry staining found that the heterogeneity in the levels of individual viral proteins was not markedly different for these synchronized infections than in the absence of pre-binding and washing.
Finally, viral mRNA expression from the secondary spread of virus from infected cells does not appreciably occur during the timeframes of our experiments, since Figure 4B does not show the pervasive presence of mixed barcodes that would occur in this case.
Therefore, variability in the timing of infection is not the dominant cause of the cell-to-cell heterogeneity in our experiments.
}
\end{quote}
}

@@

- As the variability of the viral RNA load appears to be the central result of the manuscript, it would be useful to:

    a. Employ a statistic to quantify the variability (e.g. entropy or Gini index)

    b. Use simple models to illustrate how much variability can be accounted for by simple effects, such as expected Poisson co-infection frequency, or the likelihood of attaining full complement of RNP genes.

{\color{black}
The idea of quantifying the variability in viral mRNA across infected cells is a good one.
We have done this using the Gini coefficient.
The results are in Figure 4--Figure supplement 3, and are referred to in the text related to that figure.
As a fun point of comparison, the Gini coefficients for the variability of viral mRNA across cells ($\ge$0.64 for all samples) exceeds the Gini coefficient for the unevenness of income distribution in the United States.
}

@@

{\color{red}
Still need to think about a simple model.
}

2. While the silent tagging method used to address co-infection is clever and appreciated, the issue is not fully settled. The dismissal of co-infection as a factor influencing cell-to-cell variation in lines 127-9 is based on too few cells to draw any conclusions, and thus needs to be tempered. Also, there is likely to be a significant amount of cryptic co-infection with identical barcode viruses (expected to be similar to that of mixed barcodes, ~10\%) that could influence measured heterogeneity. These points should be made in the text.

{\color{black}
These are very good points.
We have added the following text that clearly draws out the fact that our approach will only identify half the co-infections:
\begin{quote}
\textsl{
In addition, the synonymous viral barcodes only identify co-infections by viruses with different barcodes -- since the barcodes are at roughly equal proportion, we expect to miss about half of the co-infections.
Since we annotate about $\sim$10\% of the infected cells as co-infected by viruses with different barcodes (Figure 4D), we expect another $\sim$10\% of the infected cells to also be co-infected but not annotated as so by our approach.
}
\end{quote}

The reviewers are correct that our statement about the role of co-infection in influencing cell-to-cell variation was too strong.
We have clarified this statement.
We have also added a figure supplement (Figure 4-Figure supplement 5) which uses an independent method (flow cytometry staining of individual viral proteins) to support the fact that even at high infectious dose both low- and high-producing cells remain, although their relative proportions change.
This figure supplement also supports our clarified statement that viral mRNA (or protein) production is not a simple continuous function of infectious dose, but is also affected by other sources of heterogeneity.
The new text reads:
\begin{quote}
\textsl{
Notably, Figure 4E shows that there are co-infected cells with both low and high amounts of viral mRNA, suggesting that the initial infectious dose does not drive a simple continuous increase in viral transcript production.
In support of this view, we used flow cytometry to quantify the levels of individual viral proteins in cells infected at various MOIs or for which we could delineate co-infection status (Figure 4-Figure supplement 5).
This analysis shows that sub-populations of cells that express similarly low and high levels of viral proteins persist across a wide range of infectious doses, although co-infection can influence the relative proportion of infected cells that fall into these sub-populations (Figure 4-Figure supplement 5).
}
\end{quote}
}

### 2018-01-14 paper/paper.tex
added new supp fig and response related to infection timing

\subsection{Single cells show an extremely wide range of expression of viral mRNA.}
We infected A549 cells at low MOI with a mixture of the wild-type and synonymously barcoded viruses, and collected cells for sequencing at 6, 8, and 10 hours post-infection, performing two slightly different variants of the experiment for the 8-hour timepoint.
For most of the samples, we replaced the infection inoculum with fresh media at one-hour post-infection, thereby ensuring that most infection was initiated during a narrow time window.
However, for the second 8-hour sample (which we denote as ``8hr-2'' in the figures), we did \emph{not} perform this media change and instead left the cells in the original infection inoculum.
We recovered between 3,000 and 4,000 cells for each sample (Figure~\ref{fig:cells}A). 
As expected for a low-MOI infection, most cells expressed little or no viral mRNA (Figure~\ref{fig:cells}B and Figure~\ref{fig:cells}-Figure~supplement~\ref{figsupp:viral_mRNA_cumulfrac}).
Also as expected, the amount of viral mRNA per cell among infected cells increased over time (Figure~\ref{fig:cells}B and Figure~\ref{fig:cells}-Figure~supplement~\ref{figsupp:viral_mRNA_cumulfrac}).
But what was most notable was how widely the number of viral mRNA molecules varied among infected cells.
While the fraction of mRNA derived from virus was $<$0.1\% for most cells, viral mRNA constituted half the transcriptome in a few cells at 8 and 10 hours (Figure~\ref{fig:cells}B).

@@

\centerline{\includegraphics[width=0.83\linewidth]{figures/p_frac_flu_summary.pdf}}
\caption{\label{fig:viralbarcodes}
Synonymous barcodes on the viral mRNAs distinguish true infections from cells that contain viral mRNAs derived from leakage of lysed cells.
Cells with at least two viral mRNAs for which the barcode could be called, arranged in order of increasing influenza transcript counts.
Bar heights denote the number viral mRNAs on a log\textsubscript{10} scale, bar coloring is linearly proportional to the fractions of viral mRNAs derived from wild-type and synonymously barcoded virus.
Same as (A), but each bar is colored according to the relative fraction of the more common (major) and less common (minor) virus variant.
At low levels of viral mRNA there is often a roughly equal mix, suggesting contamination with viral mRNAs leaked from lysed cells.
At higher levels of viral mRNA, cells generally have only one viral variant, suggesting infection initiated by a single virion.
A few cells are also obviously co-infected with both viral variants.
We determined a threshold for calling ``true'' infections by finding the amount of viral mRNA per cell at which the viral barcode purity no longer increases with more viral mRNA.
The purity is the fraction of all viral mRNA in a cell derived from the most abundant viral barcode in that cell.
We fit a curve (orange line) to the mean purity of all cells with more than the indicated amount of viral mRNA, and drew the cutoff (dotted green line) at the point where this curve stopped increasing with the fraction of total mRNA derived from virus.
This plot illustrates the process for the 10-hour sample, see Figure~\ref{fig:viralbarcodes}-Figure~supplement~\ref{figsupp:thresholds} for similar plots for other samples.
See the Methods for details.
{\bf (D)}
The number of cells identified as infected and co-infected for each sample, as well as the number of cells with any viral read.
For all subsequent analyses, we subsampled the number of uninfected cells per sample to the greater of 50 or the number of infected cells.
{\bf (E)} 
Distribution of the fraction of mRNA per cell derived from virus for both infected and co-infected cells.
Figure~\ref{fig:viralbarcodes}-Figure~supplement~\ref{figsupp:cumulfracflu} shows these same data in a cumulative fraction plots and calculates Gini coefficients to quantify the heterogeneity in viral mRNA load.
\figsupp[Number of viral barcodes called.\label{figsupp:barcodescalled}]{
The number of viral barcodes called for each sample and gene segment. 
Viral transcripts are classified as \emph{syn} if they mapped to a synonymously barcoded influenza transcript, \emph{wt} if they mapped to a wild-type influenza transcript, \emph{invalid} if multiple reads for the same UMI differed on the status of the viral barcode, and as \emph{uncalled} if none of the reads for that UMI overlapped the region of the viral transcript containing the viral barcode.
For calculations of the number of reads in a cell derived from each viral barcode for each viral gene, the total number of detected molecules of that gene are multiplied by the fraction of those molecules with assignable barcodes that are assigned to that barcode.
}{\includegraphics[width=\linewidth]{figures/synbarcodes_umistats.jpg}}
\figsupp[Thresholds for calling infected cells. \label{figsupp:thresholds}]{
Cell lysis can lead cells to the spurious association of small amounts of extraneous mRNA with individual cells.
We wanted to avoid classifying as infected cells that had simply acquired such lysis-derived viral mRNA.
The amount of lysis-derived viral mRNA will vary among samples as a function of both the lysis rate during the cell preparation (which always varies slightly from sample to sample in the 10X procedure) and with the amount of total viral mRNA for that sample (the more viral mRNA, the more there is to be acquired from lysed cells).
As is shown in Figure~\ref{fig:viralbarcodes}B, the 8hr-2 and 10hr sample clearly have an enrichment of mixed barcodes in cells with small numbers of viral mRNA.
For each sample, we calculated the mean purity of all cells with at least the indicated amount of viral mRNA, and determined the threshold amount of viral mRNA where purity no longer increased by finding the first maxima in a loess curve fit (orange line).
We called the threshold at this point of maximum purity (dotted green line).
For the 6hr and 8hr samples there is no indication of contamination from lysis-derived reads, as Figure~\ref{fig:viralbarcodes}B shows no increase in mixed barcodes in low viral mRNA cells.
Therefore, for these samples we simply set a threshold of requiring at least 2$\times$10$^{-4}$ of the total mRNA to come from virus, which corresponds to $\sim$2 viral mRNAs for the typical cell with $10^4$ total reads (Figure~\ref{fig:cells}B).
}{\includegraphics[width=\linewidth]{figures/p_purity_threshold.pdf}}
\figsupp[Fraction of total viral mRNA from a given fraction of infected cells, and associated Gini coefficients.\label{figsupp:cumulfracflu}]{
The total fraction of all viral mRNA among infected cells that is attributable to a given fraction of these cells.
For instance, the plot for the 8hrs sample shows that $\sim$50\% of all viral mRNA is derived from $\sim$8\% of the infected cells.
The facet titles above each plot also give the Gini coefficient~\citep{gini1921measurement} that calculates the heterogeneity in the distribution of viral mRNA among infected cells.
Gini coefficients of 0 indicate a perfectly even distribution across cells, and Gini coefficients of 1 indicate a maximally skewed distribution.
}{\includegraphics[width=\linewidth]{figures/p_cumul_flu.pdf}}

\figsupp[Flow cytometry analysis of viral protein production of cells contingent on infectious dose or coinfection state. \label{figsupp:flowdose}]
{
{\bf (A)}
Expression of viral proteins in cells infected at high (MOI 5 as calculated by TCID50 on MDCK-SIAT1 cells) or low (MOI 0.1) initial infectious dose.
Cells were concurrently stained for HA and NS1 proteins 10 hours post infection.
HA staining was analyzed in cells positive for NS1 (left), and NS1 staining was analyzed in cells positive for HA (right).
While a higher dose leads to more cells expressing high amounts of viral protein, it does not greatly increase the amount of viral protein in either the low-expressing or high-expressing cells. 
Therefore, higher viral dose does not lead to a large continuous increase in viral protein production among all cells -- rather, it mostly changes the proportions of cells that fall in different parts of the highly heterogeneous distribution.
{\bf (B)}
Cells were co-infected with a mix of wild-type virus and virus in which the HA gene was replaced by GFP flanked by the terminal regions of the HA gene segment at an MOI of 0.1 for each virus. 
At 10 hours post-infection, cells were stained for NS1 and HA expression and analyzed by flow cytometry.
Cells could be annotated as infected by one or more infectious particles (wild-type or pseudovirus alone) or at least two infectious particles (coinfection).
Coinfected cells, like cells infected at a higher infectious dose, occupy different positions in the distribution of viral protein production but do not not exhibit a continuous increase in viral protein production.
Replicates shown.
}{\includegraphics[width=\linewidth]{figures/Coinf.jpg}}
\end{figure}

@@

Strikingly, the extreme variation in the number of viral transcripts per cell remains even after we apply these rigorous criteria for annotating infected cells (Figure~\ref{fig:viralbarcodes}E). 
The fraction of viral mRNA per infected cell follows a roughly exponential distribution, with many cells having few viral transcripts and a few cells having many.
At 6 and 8 hours $<$10\% of infected cells are responsible for over half the viral transcripts, while at 10 hours $<$15\% of infected cells produce over half the viral transcripts (Figure~\ref{fig:viralbarcodes}-Figure~supplement~\ref{figsupp:cumulfracflu}).
One way to quantify the heterogeneity of a distribution is to calculate the Gini coefficient~\cite{gini1921measurement}, which ranges from 0 for a completely uniform distribution, to 1 for a maximally skewed distribution.
Figure~\ref{fig:viralbarcodes}-Figure~supplement~\ref{figsupp:cumulfracflu} shows the Gini coefficients for the distribution of viral mRNA across infected cells for each sample.
The Gini coefficients are $\ge$0.64 for all samples.
As a fun point of comparison, these Gini coefficients indicate that the distribution of viral mRNA across infected cells is more uneven than the distribution of income in the United States~\citep{alvaredo2011note}.

>>>>

\subsection{Single cells show an extremely wide range of expression of viral mRNA.}
We infected A549 cells at low MOI with a mixture of the wild-type and synonymously barcoded viruses, and collected cells for sequencing at 6, 8, and 10 hours post-infection, performing two slightly different variants of the experiment for the 8-hour timepoint.
For most of the samples, we replaced the infection inoculum with fresh media at one-hour post-infection, thereby ensuring that most infection was initiated during a narrow time window.
However, for the second 8-hour sample (which we denote as ``8hr-2'' in the figures), we did \emph{not} perform this media change and instead left the cells in the original infection inoculum.
The rationale for including a sample without a media change was to determine the importance of synchronicity of the timing of infection as discussed later in this subsection.

We recovered between 3,000 and 4,000 cells for each sample (Figure~\ref{fig:cells}A). 
As expected for a low-MOI infection, most cells expressed little or no viral mRNA (Figure~\ref{fig:cells}B and Figure~\ref{fig:cells}-Figure~supplement~\ref{figsupp:viral_mRNA_cumulfrac}).
Also as expected, the amount of viral mRNA per cell among infected cells increased over time (Figure~\ref{fig:cells}B, Figure~\ref{fig:cells}-Figure~supplement~\ref{figsupp:viral_mRNA_cumulfrac}).
But what was most notable was how widely the number of viral mRNA molecules varied among infected cells.
While the fraction of mRNA derived from virus was $<$0.1\% for most cells, viral mRNA constituted half the transcriptome in a few cells at 8 and 10 hours (Figure~\ref{fig:cells}B, Figure~\ref{fig:cells}-Figure~supplement~\ref{figsupp:viral_mRNA_cumulfrac}).

@@

\begin{figure}[t!]
\centerline{\includegraphics[width=0.85\linewidth]{figures/p_frac_flu_summary.pdf}}
\caption{\label{fig:viralbarcodes}
Synonymous barcodes on the viral mRNAs distinguish true infections from cells that contain viral mRNAs derived from leakage of lysed cells.
{\bf (A)}
Cells with at least two viral mRNAs for which the barcode could be called, arranged in order of increasing influenza transcript counts.
Bar heights denote the number viral mRNAs on a log\textsubscript{10} scale, bar coloring is linearly proportional to the fractions of viral mRNAs derived from wild-type and synonymously barcoded virus.
{\bf (B)}
Same as (A), but each bar is colored according to the relative fraction of the more common (major) and less common (minor) virus variant.
At low levels of viral mRNA there is often a roughly equal mix, suggesting contamination with viral mRNAs leaked from lysed cells.
At higher levels of viral mRNA, cells generally have only one viral variant, suggesting infection initiated by a single virion.
A few cells are also obviously co-infected with both viral variants.
{\bf (C)}
We determined a threshold for calling ``true'' infections by finding the amount of viral mRNA per cell at which the viral barcode purity no longer increases with more viral mRNA.
The purity is the fraction of all viral mRNA in a cell derived from the most abundant viral barcode in that cell.
We fit a curve (orange line) to the mean purity of all cells with more than the indicated amount of viral mRNA, and drew the cutoff (dotted green line) at the point where this curve stopped increasing with the fraction of total mRNA derived from virus.
This plot illustrates the process for the 10-hour sample, see Figure~\ref{fig:viralbarcodes}-Figure~supplement~\ref{figsupp:thresholds} for similar plots for other samples.
See the Methods for details.
{\bf (D)}
The number of cells identified as infected and co-infected for each sample, as well as the number of cells with any viral read.
For all subsequent analyses, we subsampled the number of uninfected cells per sample to the greater of 50 or the number of infected cells.
{\bf (E)} 
Distribution of the fraction of mRNA per cell derived from virus for both infected and co-infected cells.
Figure~\ref{fig:viralbarcodes}-Figure~supplement~\ref{figsupp:cumulfracflu} shows these same data in a cumulative fraction plots and calculates Gini coefficients to quantify the heterogeneity in viral mRNA load.
}

\figsupp[Number of viral barcodes called.\label{figsupp:barcodescalled}]{
The number of viral barcodes called for each sample and gene segment. 
Viral transcripts are classified as \emph{syn} if they mapped to a synonymously barcoded influenza transcript, \emph{wt} if they mapped to a wild-type influenza transcript, \emph{invalid} if multiple reads for the same UMI differed on the status of the viral barcode, and as \emph{uncalled} if none of the reads for that UMI overlapped the region of the viral transcript containing the viral barcode.
For calculations of the number of reads in a cell derived from each viral barcode for each viral gene, the total number of detected molecules of that gene are multiplied by the fraction of those molecules with assignable barcodes that are assigned to that barcode.
}{\includegraphics[width=\linewidth]{figures/synbarcodes_umistats.jpg}}

\figsupp[Thresholds for calling infected cells. \label{figsupp:thresholds}]{
Cell lysis can lead cells to the spurious association of small amounts of extraneous mRNA with individual cells.
We wanted to avoid classifying as infected cells that had simply acquired such lysis-derived viral mRNA.
The amount of lysis-derived viral mRNA will vary among samples as a function of both the lysis rate during the cell preparation (which always varies slightly from sample to sample in the 10X procedure) and with the amount of total viral mRNA for that sample (the more viral mRNA, the more there is to be acquired from lysed cells).
As is shown in Figure~\ref{fig:viralbarcodes}B, the 8hr-2 and 10hr sample clearly have an enrichment of mixed barcodes in cells with small numbers of viral mRNA.
For each sample, we calculated the mean purity of all cells with at least the indicated amount of viral mRNA, and determined the threshold amount of viral mRNA where purity no longer increased by finding the first maxima in a loess curve fit (orange line).
We called the threshold at this point of maximum purity (dotted green line).
For the 6hr and 8hr samples there is no indication of contamination from lysis-derived reads, as Figure~\ref{fig:viralbarcodes}B shows no increase in mixed barcodes in low viral mRNA cells.
Therefore, for these samples we simply set a threshold of requiring at least 2$\times$10$^{-4}$ of the total mRNA to come from virus, which corresponds to $\sim$2 viral mRNAs for the typical cell with $10^4$ total reads (Figure~\ref{fig:cells}B).
}{\includegraphics[width=\linewidth]{figures/p_purity_threshold.pdf}}

\figsupp[Cumulative distributions of viral mRNA per cell and Gini coefficients.\label{figsupp:cumulfracflu}]{
The total fraction of all viral mRNA among infected cells that is attributable to a given fraction of these cells.
For instance, the plot for the 8hrs sample shows that $\sim$50\% of all viral mRNA is derived from $\sim$8\% of the infected cells.
The facet titles above each plot also give the Gini coefficient~\citep{gini1921measurement} that calculates the heterogeneity in the distribution of viral mRNA among infected cells.
Gini coefficients of 0 indicate a perfectly even distribution across cells, and Gini coefficients of 1 indicate a maximally skewed distribution.
}{\includegraphics[width=\linewidth]{figures/p_cumul_flu.pdf}}

\figsupp[Synchronization of infection does not greatly affect heterogeneity. \label{figsupp:synchronize}]
{\jdbcomment{Alistair, can you add caption}
}{\jdbcomment{Alistair, can you add add figure}}

\figsupp[Effects of infectious dose or coinfection state. \label{figsupp:flowdose}]
{
{\bf (A)}
Flow cytometry analysis of expression of viral proteins in cells infected at high (MOI 5 as calculated by TCID50 on MDCK-SIAT1 cells) or low (MOI 0.1) initial infectious dose.
Cells were concurrently stained for HA and NS1 proteins 10 hours post infection.
HA staining was analyzed in cells positive for NS1 (left), and NS1 staining was analyzed in cells positive for HA (right).
While a higher dose leads to more cells expressing high amounts of viral protein, it does not greatly increase the amount of viral protein in either the low-expressing or high-expressing cells. 
Therefore, higher viral dose does not lead to a large continuous increase in viral protein production among all cells -- rather, it mostly changes the proportions of cells that fall in different parts of the highly heterogeneous distribution.
{\bf (B)}
Cells were co-infected with a mix of wild-type virus and pseudovirus in which the HA gene was replaced by GFP flanked by the terminal regions of the HA gene segment at an MOI of 0.1 for each virus. 
At 10 hours post-infection, cells were stained for NS1 and HA expression and analyzed by flow cytometry for these proteins and GFP.
Cells could be annotated as infected by virions of the same type (wild-type infection indicated by presence of HA, or pseudovirus infection indicated by the presence of GFP) or both types of virions (indicated by presence of HA and GFP).
Coinfected cells, like cells infected at a higher infectious dose, occupy different positions in the distribution of viral protein production but do not not exhibit a continuous increase in viral protein production.
Replicates shown.
}{\includegraphics[width=\linewidth]{figures/Coinf.jpg}}
\end{figure}

@@

Strikingly, the extreme variation in the number of viral transcripts per cell remains even after we apply these rigorous criteria for annotating infected cells (Figure~\ref{fig:viralbarcodes}E). 
The fraction of viral mRNA per infected cell follows a roughly exponential distribution, with many cells having few viral transcripts and a few cells having many.
At 6 and 8 hours $<$10\% of infected cells are responsible for over half the viral transcripts, while at 10 hours $<$15\% of infected cells produce over half the viral transcripts (Figure~\ref{fig:viralbarcodes}-Figure~supplement~\ref{figsupp:cumulfracflu}).
One way to quantify the heterogeneity of a distribution is to calculate the Gini coefficient~\citep{gini1921measurement}, which ranges from 0 for a completely uniform distribution, to 1 for a maximally skewed distribution.
Figure~\ref{fig:viralbarcodes}-Figure~supplement~\ref{figsupp:cumulfracflu} shows the Gini coefficients for the distribution of viral mRNA across infected cells for each sample.
The Gini coefficients are $\ge$0.64 for all samples.
As a fun point of comparison, these Gini coefficients indicate that the distribution of viral mRNA across infected cells is more uneven than the distribution of income in the United States~\citep{alvaredo2011note}.

One possible source of heterogeneity in the amount of viral mRNA per cell is variability in the timing of infection.
If some cells are infected earlier in the experiment than others, then they might have substantially more viral mRNA.
However, several lines of evidence indicate that this is not the major cause of heterogeneity across cells.
First, the sample for which the infection inoculum was never removed (8hr-2) only shows slightly more heterogeneity than samples for which the inoculum was washed away after one hour (Figure~\ref{fig:viralbarcodes}E, Figure~\ref{fig:viralbarcodes}-Figure~supplement~\ref{figsupp:cumulfracflu}), despite the fact that the potential time window for infection is much longer in the former sample.
Second, in an independent experiment, we performed completely synchronized infections by pre-binding virus to cells on ice and then washing away unbound virus before bringing the cells to 37$^{\circ}$C~\jdbcomment{Alistair, can you add relevant citation for this method?}.
As shown in Figure~\ref{fig:viralbarcodes}-Figure~supplement~\ref{figsupp:synchronize}, flow cytometry staining found that the heterogeneity in the levels of individual viral proteins was not markedly different for these synchronized infections than in the absence of pre-binding and washing.
Finally, viral mRNA expression from the secondary spread of virus from infected cells does not appreciably occur during the timeframes of our experiments, since Figure~\ref{fig:viralbarcodes}B does not show the pervasive presence of mixed barcodes that would occur in this case.
Therefore, variability in the timing of infection is not the dominant cause of the cell-to-cell heterogeneity in our experiments.
